Triangulating iconicity: From structure mapping to guessability

### 2024-04-03 legacy/TI_01_consistency.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

A consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check. However, as we do not aim for 100% consistency, this is more for informative purposes.

@@

### Consensus version
The final result of the coding consistency is a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses. We do this this by taking the (calibrated and revised) data from one coder.

```{r create_consensus_version}

d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

@@

@@

## Adding congruency measures
We add congruency measures by combining F and M features as described in the coding scheme.

```{r congruency_1}

# note that we overwrite the old d with the new consensus version here

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d.coded %>%
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

# some mappings are mutually exclusive, so we combine them

#sanity check: are they indeed mutually exclusive? these calls should come up empty
d %>% 
  mutate(C_aspect = C_iterative + C_punctual) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) %>%
  filter(C_aspect > 1)
d %>%
  mutate(C_length = C_long + C_closure) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
  filter(C_length > 1)

d <- d %>%
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

# C_weight_ subfeatures are linked to one and the same meaning feature M_weight so they are recombined into C_magnitude
# (note: this may be subject to revision, as the subfeatures may not deserve equal weight)

d <- d %>%
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores (and some others)
# any1 = 1 if an ideophone has at least 1 congruency feature of any type
# atleast2 = if an ideophone has at least 2 congruency features of any type
# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
d <- d %>%
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3

# we can have a first look at the basic distribution of guessability by iconicity
d %>%
  group_by(C_cumulative) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_any1) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_atleast2) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_ternary) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_simple) %>%
  summarise(n=n(),score=mean.na(score))

# finally write this to XLS and CSV
write_xlsx(d,path="data\\ideophones_coded.xlsx")
write.csv(d,file="data\\ideophones_coded.csv")

>>>>

Hmm, a consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength. We can export only those columns to have a quick check.

@@

* Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.

```{r write_final}

# for now I'm taking just one coder
d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
  arrange(filename) %>%
  dplyr::select(-matches('notes'))
names(d.coded) <- gsub('_SP','',names(d.coded))

@@

write_xlsx(d.coded,path="data\\ideophones_coded.xlsx")

@@

### 2024-04-03 legacy/TI_02_congruence.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

---
title: "Triangulating iconicity: Coding analysis"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

# get consensus coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

@@

# add Z score to make scores more comparable in plots across studies
d <- d %>%
  group_by(study) %>%
  mutate(score_z = scale(score,center=T,scale=T))

# get ratings data
d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language"))

# add z score for ratings
d$rating_z <- scale(d$rating,center=T,scale=T)

# write this dataset as the fullest dataset
write.csv(d,file="data/ideophones_coded_guessed_rated.csv",fileEncoding = "UTF-8")
write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")

# also write a simplified copy for easy inspection in Excel (helpful when writing)
d %>%
  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")

@@

## Figures
Here we generate some descriptive stats and create the main figures for the paper.

```{r descriptive_stats}

d %>%
  group_by(study) %>%
  summarize(max=max(score),min=min(score),mean=mean.na(score))

@@

### Figure: Overview panel
The overview panel is created separately but we use the data from the guessability studies as a canvas.

```{r figures_paper_guessability_canvas}

@@

# blank plot can be useful as a canvas 
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)

d %>% filter(filename=="Korean_23_Motion_org") %>% View()

### Figure: Guessability by study and iconicity rating

```{r figures_paper_guessability_rating}
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
  labs(title="Guessability (both studies) by iconicity rating",
       x="iconicity rating",
       y="") +
  geom_point(size=2.0,shape=21) +
  geom_smooth(method="lm") +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

### Figure: Guessability and cumulative iconicity

```{r figures_paper_guessability_coding}

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

@@

### Figure: Ratings and cumulative iconicity

```{r ratings_and_C}

# panel A: ratings by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5) +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)

### Figure: Mappings and semantic domains

```{r figure_mappings_domains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_half_violin() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_hex() +
  NULL

@@

ggplot(data=dp, aes(x=C_cumulative,y=logodds)) +
  theme_tufte() +
  labs(title="Guessability (log odds) by cumulative iconicity",
       x="cumulative iconicity",
       y="guessability (log odds)") +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

@@

### Dotplots

@@

# how do the constituents of C_magnitude fare?
icoplot("C_magnitude")
icoplot("C_weight_voice")
icoplot("C_weight_vowel")
icoplot("C_weight_tone")

@@

# facet by language to discover crosslinguistic diversity
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

@@

## Numbers

### Coding and guessability
As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.

@@

### Semantic domains
Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.

```{r coding_semanticdomains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

We can also look in more detail at the relative differences by semantic domain by building a mixed model that predicts cumulative iconicity with domain as a predictor.

```{r coding_category_lme}

m0 <- lm(C_cumulative ~ category, data=d)
effsizes(m0)

Another prediction is that Sound and Motion ideophones should receive higher iconicity ratings than other domains. This prediction is also borne out.

```{r rating_domains}

group1 <- subset(dp.categories, category_binary == "SoundMotion")$rating
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$rating
t.test(group1,group2)
cohen.d(group1,group2)

### Coding and ratings

As above, the simplest thing to do is run a simple correlation: there is an r=0.43 correlation between meaning iconicity ratings and cumulative iconicity.

```{r coding_ratings}

cor.test(d$rating_z,d$C_cumulative,method="pearson")

As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away. However, the studies are also different in terms of

```{r coding_ratings_lme}

m0 <- lmer(rating ~ study + (1|ideophone), data=d)
m1 <- lmer(rating ~ study + C_cumulative + (1|ideophone), data=d)
anova(m0,m1)

m3 <- lmer(rating ~ study + C_aspect + (1|ideophone), data=d)
m4 <- lmer(rating ~ study + C_aspect + C_magnitude + (1|ideophone), data=d)
m5 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + (1|ideophone), data=d)
m6 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + (1|ideophone), data=d)
m7 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + C_irregular + (1|ideophone), data=d)
m8 <- lmer(rating ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)

### Control analysis: results stand even when leaving out C_modality

The `C_modality` congruence measure is intended to capture the fact that spoken words can show a range of acoustic similarities to sonic events, independently from other structural correspondences (in other words, it separates imagic iconicity from Gestalt iconicity and relative iconicity). We don't code for specific acoustic correspondences, because these would be by nature only apply to ideophones evocative of sound and would greatly complicate the coding scheme.

This makes it the least fine-grained coding category. And given that it applies to a semantic domain that is known, from prior work and first principles, to be easiest to guess, one could think of `C_modality` as just giving sound words a leg up in the cumulative scores. Therefore, as a control, it is useful to know whether the patterns we find hold up even if we disregard `C_modality` entirely.

```{r sound_off}

# we add C_nosound as a cumulative measure that excludes modality
d <- d %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude)

# recreate dp to include C_nosound
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

d %>%
  group_by(C_nosound) %>%
  summarise(n=n(),score=mean.na(score))

ggplot(data=dp, aes(x=C_nosound,y=score_z,colour=C_nosound)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity (excluding modality)",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

There is an r=0.26 correlation between logodds and C_sound across the studies. This is present in both studies.

```{r congruency_stats_nosound_1}

cor.test(d$logodds,d$C_nosound,method="pearson")

d.Collabra <- d %>% filter(study == "Collabra")
d.Language <- d %>% filter(study == "Language")
cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")
cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")

p0 <- cor.test(d$logodds,d$C_nosound,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")

The mixed effects models look much the same.
A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.

For the fixed effects, m1 lists a 0.205 improvement in logodds for `C_nosound`. According to my interpretation this means that every added iconic feature leads to an ~8% boost in proportion correct. The lower baseline for the Language set is estimated as -0.513, which translates to a 19.5% guessability penalty.

```{r congruency_stats_nosound2}

m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
m1 <- lmer(logodds ~ study + C_nosound + (1|ideophone), data=d)
anova(m0,m1)

probitlink(0.205,inverse=T)-0.5
probitlink(-0.513,inverse=T)-0.5

## Exploring coding sensitivity and congruency blind spots
What are we not capturing? Our coding and congruency measures are designed to be reproducible and human-codable. They are not exhaustive, and there may well be aspects that we are missing. Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds). Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings. We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.

@@

Coding sensitivity varies by feature, in part because not all are equally frequent. For instance, `C_weight_tone` captures ideophones that have high tone/intonation and small/light meaning or conversely low tone/intonation and big/heavy meaning.

```{r coding_sensitivity}

icoplot("C_weight_tone")

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  View()

### Variation by language

Although we have focused on cross-linguistically attested features, features may also vary by language. For instance, irregularity may work somewhat in Japanese, Korean and Ewe, the 4 Semai ideophones scored as showing this mapping have a mean accuracy score of 0.45, dragging down the average. It looks like partial reduplication really has a different function in Semai, and one that is not captured by our coding scheme.

```{r coding_by_language}

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating),C=mean.na(C_cumulative))

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  dplyr::select(ideophone,meaning,language,category,score,rating,C_irregular,C_cumulative) %>%
  View()

mS0 <- lmer(logodds ~ study + (1|ideophone), data=d)
mS1 <- lmer(logodds ~ study + language + (1|ideophone), data=d) # fails to converge
mS2 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
mS3 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + language + (1|ideophone), data=d)

anova(mS2,mS3)

fixef(mS3)

One might conclude from that language would be a relevant factor to take into account in the statistical models. But its influence is not very large. If we add it to the most complete model above, the fit is marginally improved, but the difference is not of the same magnitude as other factors. Also, language is partly confounded with quality of recordings, which are lab recordings for Ewe, Korean, Japanese and field recordings for Siwu and Semai. We can dummy code this as a new variable to check its effect.

```{r coding_by_audio}

d <- d %>%
  mutate(audio = ifelse(language %in% c("Siwu","Semai"),"field","lab"))

d %>%
  group_by(audio) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

ggplot(data=d, aes(x=audio,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="audio",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

ggplot(data=d, aes(x=language,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="language",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

# ratings look like they have a distinctly non-normal distribution for field recordings, with a large bulge below 3 (= middle)
ggplot(data=d, aes(x=audio,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

# and indeed the bulge isn't really there for any of the non-field languages (and arguably mostly attributable to the Semai recordings)
ggplot(data=d, aes(x=language,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="language",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

>>>>

---
title: "Triangulating iconicity: From structure mapping to guessability"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

@@

# get coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()

@@

# add Z score to account for differences in base success rate of Collabra and Language Studies
d <- d %>%
  group_by(study) %>%
  mutate(score_z = scale(score,center=T,scale=T))

@@

We add congruency measures that link F and M elements.

* TO DISCUSS: We need to think about the evidential value of each of the congruency measures. Most apply to many words, which is good. If some of them only capture a few (as `C_durative` and `C_weight_tone` seem to do), they are not so interesting because they don't really capture recurring iconic mappings in the data. 
* Discussed: This made us realize a need to name `M_durative` to `M_long`, covering extent not just in time but also in space (a very concrete kind of mapping for `F_finallength`.

```{r congruency_1}

# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
d <- d %>%
  mutate(C_modality = ifelse(M_sound == 1,1,0),
         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
         )

# some mappings seem to be mutually exclusive, so it might be fitting to combine them
d %>%
  mutate(C_aspect = C_iterative + C_punctual) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_event) %>%
  filter(C_aspect > 1)

d %>%
  mutate(C_length = C_long + C_closure) %>%
  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
  filter(C_length > 1)

d <- d %>%
  mutate(C_aspect = C_iterative + C_punctual,
         C_length = C_long + C_closure)

# C_weight is recombined as all three form subfeatures are linked to one and the same meaning feature M_weight

d %>% # sanity check
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone) %>%
  dplyr::select(ideophone,meaning_NL,language,C_weight_voice,C_weight_vowel,C_weight_tone,C_weight) %>%
  filter(C_weight > 0)

d <- d %>%
  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent

# add cumulative iconicity scores
d <- d %>%
  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
         C_atleast2 = ifelse(C_cumulative > 1,1,0),
         C_ternary = ifelse(C_cumulative == 0,0,
                            ifelse(C_cumulative == 1,1,
                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3

# add a simplified cumulative score by summing only the 3 correspondences that account for more than 10% of attested congruent mappings

d <- d %>%
  mutate(C_simple = C_modality + C_aspect + C_magnitude)

# write this dataset for use in the next step
write.csv(d,file="data/ideophones_coded_guessed.csv")

# also write a simplified copy for easy inspection in Excel (helpful when writing)
d %>%
  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
  write_xlsx(path="data/ideophones_coded_guessed_simplified.xlsx")

d %>%
  group_by(C_cumulative) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_any1) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_atleast2) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_ternary) %>%
  summarise(n=n(),score=mean.na(score))

d %>%
  group_by(C_simple) %>%
  summarise(n=n(),score=mean.na(score))

@@

### Figures
To visualise the results we create the figures for the paper here.

```{r figures_paper}

@@

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

@@

@@

ggplot(data=dp, aes(x=C_cumulative,y=logodds)) +
  theme_tufte() +
  labs(title="Guessability (log odds) by cumulative iconicity",
       x="cumulative iconicity",
       y="guessability (log odds)") +
  geom_half_boxplot() +
  geom_half_point() +
  NULL
ggsave("figures\\coding_guessability_by_cumulative.png",height=5,width=7.5)

@@

## Dotplots

@@

@@

d %>%
  plot_ly(x = ~logodds, 
          y = ~C_cumulative,
          type="scatter3d",
          mode="markers",
          color= ~rating_z,
          name=~ideophone,
          showlegend=F) %>%
  layout(showlegend=F)

ggsave("figures\\paper-guessability-blank.eps",height=5,width=7.5)

@@

## Statistical tests
As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.

@@

## Exploring congruency blind spots
What are we not capturing? Our coding and congruency measures are designed to be reproducible and human-codable. They are not exhaustive, and there may well be aspects that we are missing. Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds). Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings. We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.

@@
